Syntactically Informed Natural Language Inference

### 2019-03-26 paper/main.tex
Cleaning up writeup.

\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

@@

We extend the decomposable attention (DA) model from  \citet{Parikh2016-em} by
incorporating syntactic features.
The DA model obtained state-of-the-art results on the
SNLI \citep{Bowman2015-is} dataset at its time of publication while using
drastically fewer parameters than previous NLI models.
We primarily
evaluate on the domain-specific SciTail dataset \citep{Khot2018-th}
for both its smaller size and lack of annotation artifacts when compared to SNLI
\citep{Gururangan2018-lj}.

Our model, \texttt{syntail}, achieves 4.8\% better 
test accuracy than the vanilla
DA model and 1.9\% better test accuracy than DA with ELMo \citep{Peters2018-fz}
all while maintaining the core DA architecture.

@@

We experiment with features produced by two syntax parsers---the
minimal span-based neural constituency parser \citep{Stern2017-co} and
the deep biaffine attention model for dependency parsing \citep{Dozat2016-gs}. We also experiment
with different methods of incorporating the syntactic features extracted
from these parsers.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/v1.png}
    \caption{Syntail v1 architecture.}
\label{fig:v1}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/v2.png}
    \caption{Syntail v2 architecture.}
\label{fig:v2}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/v3.png}
    \caption{Syntail v3 architecture.}
\label{fig:v3}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/v4.png}
    \caption{Syntail v4 architecture.}
\label{fig:v4}
\end{figure}

@@

% \jbcomment{If we're only using data from the dependency parser we should probably 
% remove the section following this comment}
% Supplemental variations were done on top of these restructures of the network. One 
% of the variations that we introduced was a retrained constituency parser. By 
% retraining the constituency parser on the Penn Treebank dataset \citep{marcus1993building} 
% we could change the dimensions of the encoded syntax vectors to better fit with our models. 
% Additionally, we experimented with training our models using the dependency parser model 
% derived by \citet{Dozat2016-gs} to see if the different types of encoded parses influenced 
% the performance of the network.

@@

Due to resource constraints, we limited our models to only use 300-dimensional
6B GloVe word embeddings \citep{Pennington2014-uo}.
Additionally, contextual embeddings are known to capture syntax in some capacity,
so limiting our models to only use GloVe vectors
made it easier to experiment with the syntactic features in isolation.

@@

We compare our model's performance to the decomposable attention model
\citep{Parikh2016-em} and the ESIM model \citep{Chen2016-wl}.
Early on in our experiments, we noticed that features from a dependency
parser seemed to perform better than those from a constituency parser.
Therefore, all of our models use the deep biaffine attention model for
dependency parsing \citep{Dozat2016-gs}.

\paragraph{Datasets.}
For practical reasons, we only experiment on the SciTail dataset. SciTail (27k pairs) is
significantly smaller than SNLI (570k pairs) \citep{Bowman2015-is} or MultiNLI (433k pairs)
\citep{Williams2017-uh} which makes it feasible to try many different
architectures given limited resources.

Additionally, \citet{Gururangan2018-lj} showed that SNLI and MultiNLI
have significant annotation artifacts whereas SciTail is the first entailment set to be
completely derived from sentences that exist ``in the wild.''

\paragraph{Baselines.}
We list baseline performance reported by \citet{Khot2018-th} of a majority class predictor,
n-gram model, and ESIM \citep{Chen2016-wl} in table~\ref{table:experiments}. We reimplemented
and retrained the DA model and obtained similar test accuracies as reported by \citet{Chen2016-wl}.
Adding ELMo to the DA model improved test accuracy performance by approximately 3\%.

@@

\begin{table}[h]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{cccccc} 
\toprule
Model          & Embeddings      & Dev. Acc. & Test Acc.  \\ 
\midrule
Majority Class & n/a             & 63.3      & 60.3       \\
N-Gram         & n/a             & 65.0      & 70.6       \\
ESIM           & GloVe 840B 300d & 70.5      & 70.6       \\
DA             & GloVe 6B 300d   & 70.4      & 72.6       \\
DA             & ELMo            & 79.1      & 75.5       \\
\midrule
Syntail v1     & GloVe 6B 300d   & 78.4      & \textbf{77.4}       \\
Syntail v2     & GloVe 6B 300d   & \textbf{82.3}      & \textbf{77.4}       \\
Syntail v3     & GloVe 6B 300d   & 79.0         & 76.1          \\
Syntail v4     & GloVe 6B 300d   & 81.3      & 77.0       \\
\bottomrule
\end{tabular}}
\caption{Dev and test accuracies on SciTail.}
\label{table:experiments}
\end{table}

@@

\begin{table}[h]
    \centering
    \begin{tabular}{c c c}
        \toprule
         Model & Entails (correct) \\
         \midrule
         DA & 29.5\% \\
         Syntail v1 & \textbf{67.9\%} \\
         \bottomrule
    \end{tabular}
    \caption{Entailment probabilities generated by the DA and syntail v1 models on the input
             from table~\ref{table:example}.}
\label{table:pred-example}
\end{table}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/da-h2p-attention.png}
    \caption{Hypothesis to premise attention weights of the DA model when processing
             the input from table~\ref{table:example}.}
    \label{fig:da-attention-weights}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/syntail-h2p-attention.png}
    \caption{Hypothesis to premise attention weights of syntail v1 when processing
             the input from table~\ref{table:example}.}
    \label{fig:sytail-attention-weights}
\end{figure}

@@

Previous works have experimented with incorporating syntactic features for semantic tasks.
\citet{Chen2016-wl} use tree structures generated by a constituency parser to improve
their ESIM model. They encode the phrase-structure trees with a Tree-LSTM and use its
hidden states. \citet{Khot2018-th} incorporate syntactic features by
proposing a model architecture which can use any graph-based syntactic/semantic structure.

Other works have experimented with incorporating syntactic learning objectives.
\citet{Strubell2018-qw} achieved state-of-the-art performance
for a model using predicted predicates and standard word embeddings on
CoNLL-2005 SRL by performing multi-task training on additional syntactic tasks.
\citet{Swayamdipta2017-ct} use hard-coded constituency and dependency parse features
and a syntactic scaffold to
extend the softmax margin segmental RNN (segRNN) model.

@@

\bibliographystyle{acl_natbib}
\bibliography{references}

>>>>

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

@@

We extend the decomposable attention (DA) model from
\citet{parikh2016decomposable} by incorporating syntactic features.  The DA
model obtained state-of-the-art results on the SNLI \citep{bowman2015large}
dataset at its time of publication while using drastically fewer parameters
than previous NLI models.  We primarily evaluate on the domain-specific SciTail
dataset \citep{Khot2018-th} for both its smaller size and lack of annotation
artifacts when compared to SNLI \citep{gururangan2018annotation}.

Our model, \texttt{syntail}, achieves 4.8\% better test accuracy than the
vanilla DA model and 1.9\% better test accuracy than DA with ELMo
\citep{peters2018deep} all while maintaining the core DA architecture.

@@

We experiment with features produced by two syntax parsers---the minimal
span-based neural constituency parser \citep{stern2017minimal} and the deep
biaffine attention model for dependency parsing \citep{dozat2016deep}. We also
experiment with different methods of incorporating the syntactic features
extracted from these parsers.

@@

\begin{figure}[htb]
    \centering
    \includegraphics[width=\linewidth]{figures/v1.png}
    \caption{Syntail v1 architecture.}
\label{fig:v1}
\end{figure}

@@

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/v2.png}
  \caption{Syntail v2 architecture.}
\label{fig:v2}
\end{figure}

@@

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/v3.png}
  \caption{Syntail v3 architecture.}
\label{fig:v3}
\end{figure}

@@

\begin{figure}[htb]
    \centering
    \includegraphics[width=\linewidth]{figures/v4.png}
    \caption{Syntail v4 architecture.}
\label{fig:v4}
\end{figure}

@@

@@

Due to resource constraints, we limited our models to only use 300-dimensional
6B GloVe word embeddings \citep{pennington2014glove}.  Additionally, contextual
embeddings are known to capture syntax in some capacity, so limiting our models
to only use GloVe vectors made it easier to experiment with the syntactic
features in isolation.

@@

We compare our model's performance to the decomposable attention model
\citep{parikh2016decomposable} and the ESIM model \citep{chen2016enhanced}.
Early on in our experiments, we noticed that features from a dependency parser
seemed to perform better than those from a constituency parser.  Therefore, all
of our models use the deep biaffine attention model for dependency parsing
\citep{dozat2016deep}.

\paragraph{Datasets.} For practical reasons, we only experiment on the SciTail
dataset. SciTail (27k pairs) is significantly smaller than SNLI (570k pairs)
\citep{bowman2015large} or MultiNLI (433k pairs) \citep{williams2017broad}
which makes it feasible to try many different architectures given limited
resources.

Additionally, \citet{gururangan2018annotation} showed that SNLI and MultiNLI
have significant annotation artifacts whereas SciTail is the first entailment
set to be completely derived from sentences that exist ``in the wild.''

\paragraph{Baselines.}
We list baseline performance reported by \citet{Khot2018-th} of a majority
class predictor, n-gram model, and ESIM \citep{chen2016enhanced} in
table~\ref{table:experiments}. We reimplemented and retrained the DA model and
obtained similar test accuracies as reported by \citet{chen2016enhanced}.
Adding ELMo to the DA model improved test accuracy performance by approximately
3\%.

@@

\begin{table}[htb]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{cccccc} 
\toprule
Model          & Embeddings      & Dev. Acc.     & Test Acc.     \\ 
\midrule
Majority Class & n/a             & 63.3          & 60.3          \\
N-Gram         & n/a             & 65.0          & 70.6          \\
ESIM           & GloVe 840B 300d & 70.5          & 70.6          \\
DA             & GloVe 6B 300d   & 70.4          & 72.6          \\
DA             & ELMo            & 79.1          & 75.5          \\
\midrule
Syntail v1     & GloVe 6B 300d   & 78.4          & \textbf{77.4} \\
Syntail v2     & GloVe 6B 300d   & \textbf{82.3} & \textbf{77.4} \\
Syntail v3     & GloVe 6B 300d   & 79.0          & 76.1          \\
Syntail v4     & GloVe 6B 300d   & 81.3          & 77.0          \\
\bottomrule
\end{tabular}}
\caption{Dev and test accuracies on SciTail.}
\label{table:experiments}
\end{table}

@@

\begin{table}[htb]
  \centering
  \begin{tabular}{c c c}
    \toprule
    Model & Entails (correct) \\
    \midrule
    DA & 29.5\% \\
    Syntail v1 & \textbf{67.9\%} \\
    \bottomrule
  \end{tabular}
  \caption{Entailment probabilities generated by the DA and syntail v1 models
      on the input from table~\ref{table:example}.}
\label{table:pred-example}
\end{table}

@@

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/da-h2p-attention.png}
  \caption{Hypothesis to premise attention weights of the DA model when
      processing the input from table~\ref{table:example}.}
\label{fig:da-attention-weights}
\end{figure}

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/syntail-h2p-attention.png}
  \caption{Hypothesis to premise attention weights of syntail v1 when
      processing the input from table~\ref{table:example}.}
\label{fig:sytail-attention-weights}
\end{figure}

@@

Previous works have experimented with incorporating syntactic features for
semantic tasks.  \citet{chen2016enhanced} use tree structures generated by a
constituency parser to improve their ESIM model. They encode the
phrase-structure trees with a Tree-LSTM and use its hidden states.
\citet{Khot2018-th} incorporate syntactic features by proposing a model
architecture which can use any graph-based syntactic/semantic structure.

Other works have experimented with incorporating syntactic learning objectives.
\citet{strubell2018linguistically} achieved state-of-the-art performance for a
model using predicted predicates and standard word embeddings on CoNLL-2005 SRL
by performing multi-task training on additional syntactic tasks.
\citet{swayamdipta2017frame} use hard-coded constituency and dependency parse
features and a syntactic scaffold to extend the softmax margin segmental RNN
(segRNN) model.

@@

\bibliographystyle{acl_natbib_nourl}
\bibliography{references}
